### 2011-09-25 proposal/proposal.tex
Empty file for project proposal.

\documentclass{article}

\title{15-744 Project Proposal}
\author{Carlo Angiuli and Michael Sullivan}
\date{September 28, 2011}

\begin{document}
\maketitle

% Needs to include:
%  - problem statement
%  - state-of-the-art
%  - work plan / milestones

\end{document}